Exercise 3: Software Architecture in Practice Case: Beer Web Store

### 2008-03-17 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Last corrections.

\begin{description}
    \item[a. Choose the architectural drivers]
  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectural drivers. The quality attributes would be expressed as quality attribute scenarios.
    \item[b. Choose an architectural pattern]
  Based on the architectural drivers selected in a, use the mapping in Bass et al. as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system. Then select some architectural patterns to realize the tactics.
    \item[c. Instantiate modules using multiple views]
  Based on the tactics and patterns found above the architecture for the system may be expressed using UML. It is \emph{always} required to use multiple views in order to properly describe the architecture. A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint. These three viewpoints are expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively. These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
    \item[d. Define interfaces]
  Define the interfaces between the main components (the layers).
    \item[e. Verify scenarios and use cases]
  Make sure that the steps a - d covered all the parts of the scenarios. If not, include the missing parts and updates the scenarios/architecture if inconsistencies are found.
\end{description}
\subsubsection{Repeat the steps above for every module that needs further decomposition}
The first iteration only look at the overall architecture (the location and communication between layers). The following iterations will decompose the individual layers and as needed the individual components in the layers.

@@

\subsubsection{Elements}
Based on the present decomposition the elements that are know are the layers, where each layer is an element as described below.
\begin{itemize}
    \item[Clients] The clients access the server in order to perform some action.
    \item[Server] The server is accessed by the clients and when needed access the database via its Fa\c cade pattern.
    \item[Database] The database exposes its interface to the server via a Fa\c cade pattern.
\end{itemize}
The above is quickly sketched in the box-and-line drawing below in figure \ref{fig:module_view}, where the layering may also be seen. 
\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/module_view,scale=0.9 }}
\caption{Element overview}
\label{fig:module_view}
\end{figure}

\subsubsection{Relations}
Figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour. These may be expressed by e.g. a sequence diagram, as shown in Figure \ref{fig:sequence}. In this it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.

@@

There are three class of architectural elements defined.
\begin{itemize}
    \item[Processing elements] Transforms data elements.
    \item[Data elements] Information to be transformed or used.
    \item[Connecting elements] The \emph{glue} that hold the architecture together. May at times be \emph{Data elements}, \emph{Connecting elements} or both. The connecting elements are the relations between the different elements, e.g. procedure calls or message parsing.
\end{itemize}

@@

\begin{question}
Reflect on what happens if the words ``software'' and
``computing'' are removed from the definition of 'software
architecture' in [Bass et al., 2003]
\end{question}

If the references to software and computing is removed the definition talks about elements, structures and properties, and their relations. This definition may be used about anything which may be broken down into generalised components. As an example could be used the construction of a rifle. Describing the elements that goes into making a rifle (hammer, barrol, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.). The elements may fit together in different ways and multiple structures may be required to complete the description. (the hammer up, the hammer down, etc.). The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture. These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, production time, etc. while things like the cheek-pad may be optional.

The description of an architecture, as defined by Bass et. al. is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.

>>>>

\begin{description}
    \item[a. Choose the architectural drivers]
  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectural drivers. The quality attributes would be expressed as quality attribute scenarios.
    \item[b. Choose an architectural pattern]
  Based on the architectural drivers selected in \emph{a}, use the mapping in \cite{bass2003sa} as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system. Then select some architectural patterns to realize the tactics.
    \item[c. Instantiate modules using multiple views]
  Based on the tactics and patterns found above the architecture for the system may be expressed using UML. It is \emph{always} required to use multiple views in order to properly describe the architecture. A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint. These three viewpoints are expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively. These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
    \item[d. Define interfaces]
  Define the interfaces between the main components (the layers).
    \item[e. Verify scenarios and use cases]
  Make sure that the steps a - d covered all the parts of the scenarios. If not, include the missing parts and updates the scenarios/architecture if inconsistencies are found.
\end{description}
\subsubsection{Repeat the steps above for every module that needs further decomposition}
The first iteration only look at the overall architecture (the location and communication between layers). The following iterations will decompose the individual layers and as needed the individual components in the layers.

@@

\subsubsection{Elements}
Based on the present decomposition the elements that are known are the layers, where each layer is an element as described below.
\begin{itemize}
    \item[Clients] The clients access the server in order to perform some action.
    \item[Server] The server is accessed by the clients and when needed access the database via its Fa\c cade pattern.
    \item[Database] The database exposes its interface to the server via a Fa\c cade pattern.
\end{itemize}
The above is quickly sketched in the box-and-line drawing below in figure \ref{fig:module_view}, where the layering may also be seen. 
\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/module_view,scale=0.9 }}
\caption{Element overview}
\label{fig:module_view}
\end{figure}

\subsubsection{Relations}
Figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behavior. These may be expressed by e.g. a sequence diagram, as shown in Figure \ref{fig:sequence}. In this it may be seen that the Client and Server has a weak semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.

@@

There are three classes of architectural elements defined.
\begin{itemize}
    \item[Processing elements] Transforms data elements.
    \item[Data elements] Information to be transformed or used.
    \item[Connecting elements] The \emph{glue} that hold the architecture together. May at times be \emph{Data elements}, \emph{Connecting elements} or both. The connecting elements are the relations between the different elements, e.g. procedure calls or message parsing.
\end{itemize}

@@

\begin{question}
Reflect on what happens if the words ``software'' and
``computing'' are removed from the definition of 'software
architecture' in \cite{bass2003sa}
\end{question}

If the references to software and computing is removed the definition talks about elements, structures and properties, and their relations. This definition may be used about anything which may be broken down into generalised components. As an example could be used the construction of a rifle. Describing the elements that goes into making a rifle (hammer, barrel, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.). The elements may fit together in different ways and multiple structures may be required to complete the description. (the hammer up, the hammer down, etc.). The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture. These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, production time, etc. while things like the cheek-pad may be optional.

The description of an architecture, as defined by \cite{bass2003sa} is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.
